Frame Competition - Assessing the Impact of News Framing on Parties' Electoral Performance with Document Embeddings

### 2023-04-14 paper/Manuscript.Rmd
make design consistent

I test this argument using data from German newspapers 2013-2019. During the first half of this period leading up to the lower house election 2017, the immigration issue - which became salient after the 2015 refugee movements - dominated public debate in the media and among parties, as well as voters' perception of the most important problems in national politics. Using vector representations of media content and parties' manifestos (so-called document embeddings), I track the similarity of the migration discourse in national newspapers to parties' communication about migration. Combining these daily similarity estimates with polling data from *Poll of Polls*[^1] using vector auto-regression (VAR) models [@Sims1980] and an instrumental variable approach, I assess whether parties' electoral prospects increase when the media discourse is more similar to a given party's representation of the migration issue. The findings potentially challenge the current model of party competition, underlining parties' agency even when the agenda is externally determined, and speak to the real-world implications of issue framing. More generally, this research locates political opportunity spaces within debates, rather than viewing political competition solely as a game of changing issue attention.

>>>>

I test this argument using data from the German 2017 Bundestag campaign, during which the immigration issue - which became salient after the 2015 refugee movements - dominated public debate in the media and among parties, as well as voters' perception of the most important problems in national politics. Using vector representations of media content and parties' manifestos (so-called document embeddings), I track the similarity of the migration discourse in national newspapers to parties' communication about migration. Combining these daily similarity estimates with polling data from *Poll of Polls*[^1] using vector auto-regression (VAR) models [@Sims1980] and an instrumental variable approach, I assess whether parties' electoral prospects increase when the media discourse is more similar to a given party's representation of the migration issue. The findings potentially challenge the current model of party competition, underlining parties' agency even when the agenda is externally determined, and speak to the real-world implications of issue framing. More generally, this research locates political opportunity spaces within debates, rather than viewing political competition solely as a game of changing issue attention.
